LFQ: Online Learning of Per-flow Queuing Policies using Deep Reinforcement Learning

### 2020-07-03 paper/paper.tex
Update paper with Tanjas/Joachims suggestions

\usepackage[nomain, toc, acronym]{glossaries}
\usepackage{subfig}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}

\newacronym{dl}{DL}{Deep Learning}
\newacronym{rl}{RL}{Reinforcement Learning}
\newacronym{ours}{OurSolution}{OurSolution}
\newacronym{cca}{CCA}{Congestion Control Algorithm}
\newacronym{aqm}{AQM}{Active Queue Management}
\newacronym{rtt}{RTT}{Round Trip Time}
\newacronym{fq}{FQ}{Fair Queuing}
\newacronym{ml}{ML}{Machine Learning}
\newacronym{bdp}{BDP}{Bandwidth Delay Product}
\newacronym{mae}{MAE}{Mean Absolute Error}
\newacronym{mse}{MSE}{Mean Squared Error}
\newacronym{qoe}{QoE}{Quality of Experience}

@@

\begin{document}
\title{Online Learning of Per-flow Queuing Policies using Deep Reinforcement Learning}

@@

\begin{abstract}
The increasing number of different, incompatible congestion control algorithms has led to an increased deployment of fair queuing. Fair queuing isolates each network flow and gives it a separate queue. It can thus guarantee fairness for each flow even if the flows' congestion controls are not inherently fair. So far, each queue in the fair queuing system either has a fixed, static maximum size or is managed by an Active Queue Management (AQM) algorithm like CoDel. In this paper we design and implement an AQM mechanism that dynamically learns the optimal buffer size for each flow with respect to a specified utility function. We show that our Deep Learning based algorithm can correctly fingerprint each flow depending on its congestion control, delay and bandwidth and assign the optimal queue size dynamically. Besides that we also show that our mechanism can even learn completely online and can continuously adapt in deployment. Finally, we show that the computational overhead of our approach is not excessive as it can be tuned according to requirements. 
\end{abstract}

@@

The question of how to manage each flow's queue then arises: The simplest solution is using a static buffer size (queue size) for each flow. Another solution is to use an advanced \gls{aqm} mechanism like CoDel for each flow. CoDel aims to keep the queue length under a certain threshold, making sure that the queuing delay was smaller than 5\,ms in the last 100\,ms at least once. Otherwise it drops packets to decrease the queue length. Both of these approaches do not differentiate between flows: They apply the same logic to each flow irrespective of its congestion control and irrespective of the current bandwidth and \gls{rtt}. \cite{bachl_cocoa_2019} showed that this behavior leads to some flows not being able to claim the full bandwidth that they are entitled to. Other flows might achieve the full bandwidth but keep an unnecessary standing queue. The authors then proposed a mechanism that adjusts the queue of each flow based on its congestion control and showed that it works well for several common \glspl{cca}. However, their algorithm has parameters that have to be manually adjusted and it is not guaranteed to work for every \gls{cca} because it makes the assumption that each flow's congestion window follows a zigzag pattern. Furthermore, their approach is only tailored towards flows that always have data to send while the behavior for application limited flows, that also have idle periods, is not clear.

@@

To make our work reproducible and to encourage further research, we make the code, figures and trained neural network weights of this work publicly available\footnote{\url{https://github.com/CN-TU/reinforcement-learning-for-per-flow-buffer-sizing}}.

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{figures/cocoa_illustration_too_little.pdf}
\caption{If the buffer is too small, loss-based \glspl{cca} cannot fully utilize the link since they send too few data following the multiplicative decrease that occurs after packet loss.}
\label{fig:tooLittle}
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{figures/cocoa_illustration_too_much.pdf}
\caption{If the buffer is too large, loss-based \glspl{cca} keep an unnecessary standing queue not required for achieving full link utilization.}
\label{fig:tooMuch}
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{figures/cocoa_illustration_perfect.pdf}
\caption{For a flow controlled by New Reno, if the buffer is the same size as the \gls{bdp}, no standing queue exists and the full bandwidth is achieved.}
\label{fig:perfect}
\end{figure}

@@

For the popular \gls{cca} \textit{New Reno}, the buffer size that is necessary in a fair queuing setting is one \gls{bdp}, meaning $\textit{speed}\times\textit{delay}$ (\autoref{fig:perfect}). This is because its multiplicative decrease factor is $0.5$, meaning that after packet loss, the congestion window is halved. For other \glspl{cca} like \textit{Cubic} \cite{ha_cubic:_2008}, the multiplicative decrease factor is $0.7$, meaning that the congestion window is reduced to 70\% of its previous value upon packet loss. This means that the minimum buffer size required to achieve full throughput for a Cubic flow is $\left(\frac{1}{0.7}-1\right)\times \textit{BDP}$, which is around 43\% of the \gls{bdp} and thus less than for New Reno. It follows that an adaptive fair \gls{aqm} must learn the available bandwidth, \gls{rtt} and the underlying \gls{cca} of a flow to adjust the buffer optimally.

@@

In this utility function, the choosable parameter $\alpha$ specifies the tradeoff that is chosen between the bandwidth and the buffer size. With $\alpha$ going to zero, the optimum policy approaches the one in which the buffer size is large enough that a flow never underutilizes the link but at the same time the buffer is never going to be larger than necessary if this doesn't provide a benefit in throughput.

Our \gls{ml} system uses the above utility function to learn the optimal behavior. \gls{ours} computes features for each flow after every packet it receives and then outputs the optimal buffer size of which it thinks that it maximizes the utility function. As the input we use the following features, which are computed continuously when receiving/sending a packet: 
\begin{itemize}
\item queue size
\item standard deviation of the queue size 
\item maximimum allowed buffer size
\item rate of incoming data
\item rate of outgoing data
\item time since the last packet loss
\end{itemize}
We do not use these features directly but instead use 10 exponentially weighted averages of them for each feature with weights of $2^{-4}$, $2^{-5}$, ... , $2^{-13}$. The advantage of using exponentially weighted averages is that they do not occupy any space in memory except for their own values. This is opposed to regular moving averages, which have to keep the entire window of data in memory.

One difficulty is to get an exponentially moving average for the rate of incoming/outgoing packets because computing the average of rates is mathematically not easily possible: Instead we compute the exponentially moving average of the interarrival times of incoming packets and the interdeparture times of outgoing packets. We then invert this number to get the rate. Another issue here is that numeric instability and division by zero errors can occur using this approach. Thus, we only use the exponential average with a weight of $2^{-n}$ after at least $n$ packets have arrived. Otherwise we set it to zero. We did this because we saw that using the averages of interarrival/interdeparture times led to slower learning than using the rates. This is because if only the interarrival/interdeparture times are given, the neural network also has to learn the additional step of inverting them, making the learning more time-consuming.

@@

For the simulations we randomly draw a bandwidth (5 to 25\,Mbit/s), a delay (5 to 25\,ms) and duration (3.75\,ms to 6.25\,ms) and a \gls{cca} (New Reno or BIC \cite{lisong_xu_binary_2004}). We use New Reno because it is one of the oldest, most widely deployed \glspl{cca} and BIC because it is similar to Cubic, which is the default in Linux, Windows and most other OSs currently. We cannot use Cubic itself because there is no stable Cubic implementation in ns-3 currently. As the experiment time (the time at which the experiment is launched to determine if a larger or smaller buffer size would be better), we use a random number between 0 and half of the flow duration of the current flow.

@@

One major issue that we encountered bufferbloat (over 100\,ms even if we set our buffers to 1 packet) and we couldn't pinpoint the location where it was introduced first. The reason was that ns-3 also has a additional buffer for each link: First, there's the buffer of the queuing discipline (for which we implement our mechanism) and then there's also a hardware buffer after that. This buffer has a default queue size of 100 packets. Thus, on a 10\,Mbit/s link, this hardware buffer alone results in a delay of 120\,ms when simulating bulk transfers. We reduced this buffer to 1 packet and then found the simulations to behave as expected.

@@

The training procedure involves the following steps:
\begin{enumerate}
\item Draw a random sample of bandwidths, delays, \gls{cca} and flow durations.
\item Simulate each flow concurrently, compute the feature vector continuously and let the neural network output its optimal buffer size each time a new packet arrives. 
\item During each flow, at a random time between 0 and the flow length in seconds divided by 2, perform the experiment: Continue one simulation with the current buffer size $+1$ packet and one with the current buffer size $-1$ packet until the end of the flow. 
\item Wait until all flows are finished
\item Check for each flow which of the two versions performed better ($+1$ or $-1$) regarding the utility function.
\item Update the neural network so that it learns to output the better buffer size when being fed the inputs as they were at the time at which the experimentation started. Specifically, we compute the \gls{mae} between the output of the neural network and the desired output (the one that performed better in the experiment).
\item Start the next iteration. 
\end{enumerate}

During deployment, the trained neural network is used but the experiment step is skipped. We chose to run 20 simulations concurrently since our computers have 40 CPUs and every simulation is split in two at the time of the experiment. Thus with 20 simulations we fully utilize the 20 CPUs. This implies that the batch size used for the \gls{dl} is 20.

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{/mnt/cluster/results/RLQueueDisc/logs/plots/2020-6-6-10-24-18_prediction.png}
\caption{The prediction of the optimal buffer by the neural network during offline training with the tradeoff $\alpha=0.01$.}
\label{fig:offlineTraining}
\end{figure}

\begin{figure*}[h]
\centering
\subfloat[New Reno, varying bandwidth.\label{fig:SmallAlphaNewRenoBandwidth}
]{
\includegraphics[width=0.98\columnwidth]{{"../ns-allinone-3.30.1/ns-3.30.1/results/RLQueueDisc/logs/plots/2020-6-6-10-24-18_81000.weights_New Reno_bandwidth"}.pdf}
}{}
\subfloat[New Reno, varying delay.\label{fig:SmallAlphaNewRenoDelay}
]{
\includegraphics[width=0.98\columnwidth]{{"../ns-allinone-3.30.1/ns-3.30.1/results/RLQueueDisc/logs/plots/2020-6-6-10-24-18_81000.weights_New Reno_delay"}.pdf}
}{}
\subfloat[BIC, varying bandwidth\label{fig:SmallAlphaBicBandwidth}
]{
\includegraphics[width=0.98\columnwidth]{{"../ns-allinone-3.30.1/ns-3.30.1/results/RLQueueDisc/logs/plots/2020-6-6-10-24-18_81000.weights_Bic_bandwidth"}.pdf}
}{}
\subfloat[BIC, varying delay\label{fig:SmallAlphaBicDelay}
]{
\includegraphics[width=0.98\columnwidth]{{"../ns-allinone-3.30.1/ns-3.30.1/results/RLQueueDisc/logs/plots/2020-6-6-10-24-18_81000.weights_Bic_delay"}.pdf}
}{}
\caption{Change of the buffer size after offline learning when varying bandwidth/delay/\gls{cca}, keeping the other parameters constant at their mean. }
\label{fig:offlineSmallAlpha}
\end{figure*}

When performing offline learning with the tradeoff parameter $\alpha=0.01$ (\autoref{fig:offlineTraining}), the neural network first starts with a prediction that is close to 0 (due to the initialization of neural network weights). During the first couple of thousand training flows, it increases its output and approaches what appears to be the buffer size that works best on average (slightly higher than 15). Then the neural network starts differentiating different flow and learns custom policies for them. After around 250000 flows, the output appears to be stable and doesn't change anymore. There's one noticeable outlier at around 120000 flows. This appears to be some instability that can commonly occur in the early stages of neural network training.

\begin{figure}[h]
\includegraphics[width=\columnwidth]{{"../ns-allinone-3.30.1/ns-3.30.1/results/RLQueueDisc/queueTraces/2020-6-6-10-24-18_81000.weights/cc_0_bw_25.000000_delay_15.000000"}.pdf}
\caption{Showing the maximum queue length and queue length of a New Reno flow controlled by \gls{ours} ($\alpha=0.01$) at the bottleneck with a 25\,Mbit/s link and a delay of 15\,ms. The queue never becomes empty and also never forms a standing queue.}
\label{fig:exampleRenoLargeBw}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{{"../ns-allinone-3.30.1/ns-3.30.1/results/RLQueueDisc/queueTraces/2020-6-6-10-24-18_81000.weights/cc_0_bw_6.010101_delay_15.000000"}.pdf}
\caption{Showing the maximum queue length and queue length of a New Reno flow controlled by \gls{ours} ($\alpha=0.01$) at the bottleneck with a 6\,Mbit/s link and a delay of 15\,ms. The queue never becomes empty and also never forms a standing queue.}
\label{fig:exampleRenoSmallBw}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{{"../ns-allinone-3.30.1/ns-3.30.1/results/RLQueueDisc/queueTraces/2020-6-6-10-24-18_81000.weights/cc_1_bw_25.000000_delay_15.000000"}.pdf}
\caption{Showing the maximum queue length and queue length of a BIC flow controlled by \gls{ours} ($\alpha=0.01$) at the bottleneck with a 25\,Mbit/s link and a delay of 15\,ms. The queue never becomes empty and also never forms a standing queue.}
\label{fig:exampleBicLargeBw}
\end{figure}

@@

To have a quantifiable measure of the success of \gls{ours}, we compute the correlation between a change of bandwidth/delay and the maximum queue size output by the neural network. If the neural network learned ideally, the correlation obtained would be 1 (100\%) as, as mentioned in \autoref{sec:concept}, with a larger delay/bandwidth, a larger buffer is necessary. \autoref{tab:corrOfflineSmallAlpha} shows that the for New Reno the correlation is very high while for BIC it is high but not as good as for New Reno. We attribute this to the fact that for BIC the buffer that is required is generally smaller (only a couple of packets) and thus fluctuations in the training process are more pronounced.

@@

The results also show that increasing the delay increases the output of the neural network. This is more interesting than for the bandwidth since no input feature directly encodes the delay. Thus, the neural network must observe the pattern of the \gls{cca}, estimate the \gls{rtt} and output the suitable maximum queue size. The right column of \autoref{tab:corrOfflineSmallAlpha} shows that there is clearly a linear relationship between the delay and the output maximum queue size.

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{/mnt/cluster/results/RLQueueDisc/logs/plots/2020-6-4-15-39-25_prediction.png}
\caption{The prediction of the optimal buffer by the neural network during offline training with the tradeoff $\alpha=10$.}
\label{fig:offlineTrainingLargeAlpha}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{{"../ns-allinone-3.30.1/ns-3.30.1/results/RLQueueDisc/queueTraces/2020-6-4-15-39-25_108000.weights/cc_0_bw_6.010101_delay_15.000000"}.pdf}
\caption{Showing the maximum queue length and queue length of a New Reno flow controlled by \gls{ours} ($\alpha=10$) at the bottleneck with a 6\,Mbit/s link and a delay of 15\,ms. The queue occasionally becomes empty which is intended as with the higher tradeoff $\alpha$, \gls{ours} tries to favor a small queue more than high throughput.}
\label{fig:exampleRenoSmallBwLargeAlpha}
\end{figure}

@@

For online training we need two neural networks: 
\begin{itemize}
\item An actor network, which outputs the optimal buffer size at each time step.
\item A critic network, which outputs the utility that it predicts is going to be achieved when keeping the current buffer size until the end of the flow. 
\end{itemize}
Both of these neural networks have the same number of layers and neurons.

The training procedure involves the following steps:
\begin{enumerate}
\item Draw a random sample of bandwidths, delays, \gls{cca} and flow durations.
\item Simulate each flow concurrently, compute the feature vector continuously and let the neural network output its optimal buffer size each time a new packet arrives. 
\item During each flow, at a random time between 0 and the flow length in seconds divided by 2, perform the experiment: Continue the simulation either with the current buffer size incremented by 1 or decremented by 1 until the end of the flow. 
\item Wait until all flows are finished
\item Check for each flow if its decision was better than the expectation of the critic network.
\item Update the actor neural network to output the better buffer size when being fed the inputs as they were at the time at which the experimentation started. If the experiment was with buffer size $-1$ but this yielded worse results than expected, this implies that the buffer size $+1$ would have been better. As for offline training we use the \gls{mae}. 
\item Update the critic neural network by using the input vector that was recorded at the time at which the experiment started and the utility that was achieved as the label. Specifically, \gls{mse} is used as the loss function. 
\item Start the next iteration. 
\end{enumerate}

>>>>

\usepackage[nomain, toc, acronym]{glossaries}
\usepackage{subfig}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{float}

\newacronym{dl}{DL}{Deep Learning}
\newacronym{rl}{RL}{Reinforcement Learning}
\newacronym{ours}{LFQ}{Fair Learning Qdisc}
\newacronym{cca}{CCA}{Congestion Control Algorithm}
\newacronym{aqm}{AQM}{Active Queue Management}
\newacronym{rtt}{RTT}{Round Trip Time}
\newacronym{fq}{FQ}{Fair Queuing}
\newacronym{ml}{ML}{Machine Learning}
\newacronym{bdp}{BDP}{Bandwidth Delay Product}
\newacronym{mae}{MAE}{Mean Absolute Error}
\newacronym{mse}{MSE}{Mean Squared Error}
\newacronym{qoe}{QoE}{Quality of Experience}

@@

\begin{document}
\title{LFQ: Online Learning of Per-flow Queuing Policies using Deep Reinforcement Learning}

@@

\begin{abstract}
The increasing number of different, incompatible congestion control algorithms has led to an increased deployment of fair queuing. Fair queuing isolates each network flow and can thus guarantee fairness for each flow even if the flows' congestion controls are not inherently fair. So far, each queue in the fair queuing system either has a fixed, static maximum size or is managed by an Active Queue Management (AQM) algorithm like CoDel. In this paper we design an AQM mechanism (Fair Learning Qdisc (LFQ)) that dynamically learns the optimal buffer size for each flow according to a specified utility function online. We show that our Deep Learning based algorithm can dynamically assign the optimal queue size to each flow depending on its congestion control, delay and bandwidth. Comparing to competing fair AQM schedulers, it provides significantly smaller queues while achieving the same or higher throughput. 
\end{abstract}

@@

Given the high throughput and low queuing delay that is required by emerging cloud gaming \cite{jarschel_evaluation_2011} and virtual reality \cite{elbamby_toward_2018} applications, the optimal management of each flow's queue becomes an increasingly relevant question. Furthermore, for 5G, ultra low latency communications have been made a priority \cite{li_5g_2018} and we argue that effective buffering must also be taken into account for achieving this goal. The simplest solution to managing a flow's buffer is to a static buffer size (queue size) for each flow. However, this can lead to inacceptably large queuing delays. Another solution is to use an advanced \gls{aqm} mechanism like CoDel for each flow. CoDel aims to keep the queue length under a certain threshold, making sure that the queuing delay was smaller than 5\,ms in the last 100\,ms at least once. Otherwise it drops packets to decrease the queue length. Both of these approaches do not differentiate between flows: They apply the same logic to each flow irrespective of its congestion control and irrespective of the current bandwidth and \gls{rtt}. \cite{bachl_cocoa_2019} showed that this behavior leads to some flows not being able to claim the full bandwidth that they are entitled to. Other flows might achieve the full bandwidth but keep an unnecessary standing queue. The authors then proposed a mechanism that adjusts the queue of each flow based on its congestion control and showed that it works well for several common \glspl{cca}. However, their algorithm has parameters that have to be manually adjusted and it is not guaranteed to work for every \gls{cca} because it makes the assumption that each flow's congestion window follows a zigzag pattern. Furthermore, their approach is only tailored towards flows that always have data to send while the behavior for application limited flows, that also have idle periods, is not clear.

@@

To make our work reproducible and to encourage further research, we make the code, figures and trained neural network weights of this work publicly available\footnote{\scriptsize\url{https://github.com/CN-TU/reinforcement-learning-for-per-flow-buffer-sizing}}.

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{figures/cocoa_illustration_too_little.pdf}
\caption{If the buffer is too small, loss-based \glspl{cca} cannot fully utilize the link since they send too few data (less than the \gls{bdp}) following the multiplicative decrease that occurs after packet loss.}
\label{fig:tooLittle}
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{figures/cocoa_illustration_too_much.pdf}
\caption{If the buffer is too large, loss-based \glspl{cca} keep an unnecessary standing queue (data in flight always more than \gls{bdp}) not required for achieving full link utilization.}
\label{fig:tooMuch}
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{figures/cocoa_illustration_perfect.pdf}
\caption{For a flow controlled by New Reno, if the buffer is the same size as the \gls{bdp}, no standing queue exists and the full bandwidth is achieved.}
\label{fig:perfect}
\end{figure}

@@

For the popular \gls{cca} \textit{New Reno}, the buffer size that is necessary in a fair queuing setting is one \gls{bdp}, meaning $\textit{speed}\times\textit{delay}$ (\autoref{fig:perfect}). This is because its multiplicative decrease factor is $0.5$, meaning that after packet loss, the congestion window is halved. For other \glspl{cca} like \textit{Cubic} \cite{ha_cubic:_2008}, the multiplicative decrease factor is $0.7$, meaning that the congestion window is reduced to 70\% of its previous value upon packet loss. This means that the minimum buffer size required to achieve full throughput for a Cubic flow is $\left(\frac{1}{0.7}-1\right)\times \textit{BDP}$, which is around 43\% of the \gls{bdp} and thus less than for New Reno. That means the minimum required buffer size depends on the \gls{cca}. To reduce packet loss and standing queues we propose an adaptive fair \gls{aqm} that learns the available bandwidth, \gls{rtt} and the underlying \gls{cca} of a flow to adjust the buffer optimally.

@@

In this utility function, the choosable parameter $\alpha$ specifies the tradeoff that is chosen between the bandwidth and the buffer size. With $\alpha$ going to zero, the optimum policy approaches the one in which the buffer size is large enough that a flow never underutilizes the link but at the same time the buffer is never going to be larger than necessary if this doesn't provide a benefit in throughput. Alternatively, the delay could be used instead of the queue size because it is more closely related to \gls{qoe}. However, the delay that is caused by a certain queue length depends on the current link speed. This means that using delay would make the utility function more variable, depending on the current link speed.

Our \gls{ml} system uses the above utility function to learn the optimal behavior. \gls{ours} computes features for each flow after every packet it receives and then outputs the optimal buffer size of which it thinks that it maximizes the utility function. As the input we use the following features, which are updated continuously when receiving/sending a packet: 
\begin{itemize}
\item queue size
\item standard deviation of the queue size 
\item maximimum allowed buffer size
\item rate of incoming data
\item rate of outgoing data
\item time since the last packet loss
\end{itemize}
We do not use these features directly because this could result in large, spontaneous variation in the input features. For example, two packets might not be received in regular intervals but in bursts. This means that without any smoothening, large instability would occur. Instead use 10 exponentially weighted averages of them for each feature with weights of $2^{-4}$, $2^{-5}$, ... , $2^{-13}$. The advantage of using exponentially weighted averages is that they do not occupy any space in memory except for their own values. This is opposed to regular moving averages, which have to keep the entire window of data in memory. For example, for a regular weighted average of window size 1000, the last 1000 data points have to be kept in memory at any time, leading to a large consumption of memory.

One difficulty is to get an exponentially moving average for the rate of incoming/outgoing packets. Instead we compute the exponentially moving average of the interarrival times of incoming packets and the interdeparture times of outgoing packets. We then invert this number to get the rate. An issue with this approach is that numeric instability and division by zero errors can occur using this approach. Thus, we only use the exponential average with a weight of $2^{-n}$ after at least $n$ packets have arrived. Otherwise we set it to zero. We did this because we saw that using the averages of interarrival/interdeparture times led to slower learning than using the rates. This is because if only the interarrival/interdeparture times are given, the neural network also has to learn the additional step of inverting them, making the learning more time-consuming.

@@

For the simulations we randomly draw a bandwidth (5 to 25\,Mbit/s), a delay (5 to 25\,ms) and duration (3.75\,ms to 6.25\,ms) and a \gls{cca} (New Reno or BIC \cite{lisong_xu_binary_2004}). We chose these ranges bandwidths and delays so that the neural network encounters flows with vastly different characteristics and has to learn to deal with them. We use New Reno because it is one of the oldest, most widely deployed \glspl{cca} and BIC because it is similar to Cubic, which is the default in Linux, Windows and most other OSs currently. We cannot use Cubic itself because there is no stable Cubic implementation in ns-3 currently.

As the experiment time (the time at which the experiment is launched to determine if a larger or smaller buffer size would be better), we use a random number between 0 and half of the flow duration of the current flow. The number must be random so that the \gls{rl} learn to take decisions for every stage of a flow: The optimal maximum buffer size is different in the beginning of the flow, when you don't know any information about the flow yet, than it is after several seconds. The expected behavior is that in the beginning of a flow, the \gls{rl} simply uses the average buffer size that works the best on average for all flows. Then, after more information about the flow is known and it can be fingerprinted effectively, the estimation of the \gls{rl} is supposed to become more fine-grained and specific to the current flow. That means that the \gls{rl} has to learn what the optimal buffer size is for every stage of a flow and that's why the time at which the experiment is performed must be random. Another issue is that if, for example, the experiment is performed just before the end of the flow, the impact of the decision of the \gls{rl} is not clear: If the \gls{rl}, for example, outputs a maximum buffer size of 15 just before the end of the flow, then it has no influence on the flow because the flow ends anyway. That's why we don't perform experiments for the \gls{rl} training in the latter half of a flow.

@@

@@

The training procedure involves the following steps:
\begin{enumerate}
\item Draw a random sample of 20 bandwidths, delays, \gls{cca} and flow durations.
\item Simulate each flow concurrently, compute the feature vector continuously and let the neural network output its optimal buffer size each time a new packet arrives. 
\item During each flow, at a random time between 0 and half the flow length in seconds, perform the experiment: Fork the simulation process for each flow and continue one simulation with the current buffer size $+1$ packet and one with the current buffer size $-1$ packet until the end of the flow. This procedure is essentially \textit{A/B testing}.  
\item Wait until all flows are finished
\item Check for each flow which of the two versions performed better ($+1$ or $-1$) regarding the utility function.
\item Update the neural network so that it learns to output the better buffer size when being fed the inputs as they were at the time at which the experimentation started. Specifically, we compute the \gls{mae} between the output of the neural network and the desired output (the one that performed better in the experiment) for the whole batch of 20 results.
\item Start the next iteration. 
\end{enumerate}

During deployment, the trained neural network is used but the experiment step is skipped. We chose to run 20 simulations concurrently since our computers have 40 CPUs and every simulation is split in two at the time of the experiment when each simulation process is forked. Thus with 20 simulations we fully utilize the 20 CPUs. This implies that the batch size used for the \gls{dl} is 20.

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{{"figures/_mnt_cluster_results_RLQueueDisc_logs_plots_2020-6-6-10-24-18_prediction"}.png}
\caption{The prediction of the optimal buffer by the neural network during offline training with the tradeoff $\alpha=0.01$. The large variance in predicted optimal buffer sizes is the intended result as this means that different optimal buffer sizes are output for different types of flows, which means that the neural network can differentiate different flows.}
\label{fig:offlineTraining}
\end{figure}

\begin{figure*}[h]
\centering
\subfloat[New Reno, varying bandwidth.\label{fig:SmallAlphaNewRenoBandwidth}
]{
\includegraphics[width=0.98\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_logs_plots_2020-6-6-10-24-18_81000.weights_New Reno_bandwidth"}.pdf}
}{}
\subfloat[New Reno, varying delay.\label{fig:SmallAlphaNewRenoDelay}
]{
\includegraphics[width=0.98\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_logs_plots_2020-6-6-10-24-18_81000.weights_New Reno_delay"}.pdf}
}{}
\subfloat[BIC, varying bandwidth\label{fig:SmallAlphaBicBandwidth}
]{
\includegraphics[width=0.98\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_logs_plots_2020-6-6-10-24-18_81000.weights_Bic_bandwidth"}.pdf}
}{}
\subfloat[BIC, varying delay\label{fig:SmallAlphaBicDelay}
]{
\includegraphics[width=0.98\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_logs_plots_2020-6-6-10-24-18_81000.weights_Bic_delay"}.pdf}
}{}
\caption{Change of the buffer size after offline learning when varying bandwidth/delay/\gls{cca}. When bandwidth is varied, delay is kept in the middle of its range, at 15\,ms. Vice versa, when delay is varied, bandwidth is kept in the middle of its range, at 15\,Mbit/s.}
\label{fig:offlineSmallAlpha}
\end{figure*}

When performing offline learning with the tradeoff parameter $\alpha=0.01$ (\autoref{fig:offlineTraining}), the neural network first starts with a prediction that is close to 0 (due to the initialization of neural network weights). During the first couple of thousand training flows, it increases its output and approaches what appears to be the buffer size that works best on average (slightly higher than 15). Then the neural network gradually starts understanding and differentiating different flows and learns custom policies for them. After around 250000 flows, the output appears to be stable and doesn't change anymore. There's one noticeable outlier at around 120000 flows. This appears to be some instability that can commonly occur in the early stages of neural network training but what has no significance for the training process since it resumes normally after the outlier.

\begin{figure}[h]
\includegraphics[width=\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_queueTraces_2020-6-6-10-24-18_81000.weights_cc_0_bw_25.000000_delay_15.000000"}.pdf}
\caption{Showing the maximum queue length and queue length of a New Reno flow controlled by \gls{ours} ($\alpha=0.01$) at the bottleneck with a 25\,Mbit/s link and a delay of 15\,ms. The queue never becomes empty and also never forms a standing queue.}
\label{fig:exampleRenoLargeBw}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_queueTraces_2020-6-6-10-24-18_81000.weights_cc_0_bw_6.010101_delay_15.000000"}.pdf}
\caption{Showing the maximum queue length and queue length of a New Reno flow controlled by \gls{ours} ($\alpha=0.01$) at the bottleneck with a 6\,Mbit/s link and a delay of 15\,ms. The queue never becomes empty and also never forms a standing queue.}
\label{fig:exampleRenoSmallBw}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_queueTraces_2020-6-6-10-24-18_81000.weights_cc_1_bw_25.000000_delay_15.000000"}.pdf}
\caption{Showing the maximum queue length and queue length of a BIC flow controlled by \gls{ours} ($\alpha=0.01$) at the bottleneck with a 25\,Mbit/s link and a delay of 15\,ms. The queue never becomes empty and also never forms a standing queue.}
\label{fig:exampleBicLargeBw}
\end{figure}

@@

To have a quantifiable measure of the success of \gls{ours}, we compute the correlation between a change of bandwidth/delay and the maximum queue size output by the neural network. If the neural network learned ideally, the correlation obtained would be 1 (100\%) as, as mentioned in \autoref{sec:concept}, with a larger delay/bandwidth the \gls{bdp} is larger and a larger buffer is necessary. \autoref{tab:corrOfflineSmallAlpha} shows that the for New Reno the correlation is very high while for BIC it is high but not as good as for New Reno. We attribute this to the fact that for BIC the buffer that is required is generally smaller (only a couple of packets) and thus fluctuations in the training process are more pronounced.

@@

The results also show that increasing the delay increases the output (maximum buffer size) of the neural network. This is more interesting than for the bandwidth since no input feature directly encodes the delay. Thus, the neural network must observe the pattern of the \gls{cca}, estimate the \gls{rtt} and output the suitable maximum queue size. The right column of \autoref{tab:corrOfflineSmallAlpha} shows that there is clearly a linear relationship between the delay and the output maximum queue size as the correlation is close to 1 (or 100\%).

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{{"figures/_mnt_cluster_results_RLQueueDisc_logs_plots_2020-6-4-15-39-25_prediction"}.png}
\caption{The prediction of the optimal buffer by the neural network during offline training with the tradeoff $\alpha=10$.}
\label{fig:offlineTrainingLargeAlpha}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_queueTraces_2020-6-4-15-39-25_108000.weights_cc_0_bw_6.010101_delay_15.000000"}.pdf}
\caption{Showing the maximum queue length and queue length of a New Reno flow controlled by \gls{ours} ($\alpha=10$) at the bottleneck with a 6\,Mbit/s link and a delay of 15\,ms. The queue occasionally becomes empty which is intended as with the higher tradeoff $\alpha$, \gls{ours} tries to favor a small queue more than high throughput.}
\label{fig:exampleRenoSmallBwLargeAlpha}
\end{figure}

@@

The major change is that for online learning, A/B testing, like for the offline learning, is not possible: In the real world, each time the buffer size is updated for a flow, this is a definite decision. It is not possible to artificially ``split'' a flow like in the simulator and check if a larger or a smaller buffer would have been a better choice. Another analogy for this is that for online learning, we have to unveil what happened to Schrödinger's cat (if it is dead or not), when we launch the experiment, which buffer size is better. For the offline learning, on the other side, we can try out both versions in the simulator and continue one reality with the cat being dead (maximum buffer size $-1$) and the other one with the cat being alive (maximum buffer size $+1$).

To accommodate for this difference from offline learning, for online training we need two neural networks: 
\begin{itemize}
\item An actor network, which outputs the optimal buffer size at each time step.
\item A critic network, which outputs the utility that it predicts is going to be achieved when keeping the current buffer size until the end of the flow. 
\end{itemize}
Both of these neural networks have the same number of layers and neurons.

The training procedure involves the following steps:
\begin{enumerate}
\item Draw a random sample of 40 bandwidths, delays, \gls{cca} and flow durations.
\item Simulate each flow concurrently, compute the feature vector continuously and let the neural network output its optimal buffer size each time a new packet arrives. 
\item During each flow, at a random time between 0 and the flow length in seconds divided by 2, perform the experiment: Continue the simulation either with the current buffer size incremented by 1 or decremented by 1 until the end of the flow.
\item Wait until all flows are finished
\item Check for each flow if its decision was better than the expectation of the critic network.
\item Update the actor neural network to output the better buffer size when being fed the inputs as they were at the time at which the experimentation started. If the experiment was with buffer size $-1$ but this yielded worse results than expected, this implies that the buffer size $+1$ would have been better. As for offline training we use the \gls{mae}. 
\item Update the critic neural network by using the input vector that was recorded at the time at which the experiment started and the utility that was achieved as the label. Specifically, \gls{mse} is used as the loss function. 
\item Start the next iteration. 
\end{enumerate}
